---
title: "Homework assignment 5"
format: html
---

## Load required libraries and data

```{r}
library(tidyverse)
library(brms)
```

```{r}
hs_dep <- read_csv2("data_files/hate_depres.csv")
hs_dep <- hs_dep %>% 
  mutate(familySup = ifelse(famAcc == 0, "no", "yes"))
hs_dep
```

The dataset contains data from N = 3,827 gay and lesbian participants.

hs – Measures how often participants encountered hate speech against gay or lesbian individuals (1 = very rarely, 7 = very often).
depres – Measures the intensity of depression symptoms using the PHQ-8 scale (1 = low intensity, 5 = high intensity).
famAcc – Indicates the number of family members who accept the participant’s sexual orientation (0 = no one, 4 = both parents and siblings).
familySup – A categorical variable indicating whether the participant has at least one family member who accepts their sexual orientation ("yes" or "no").

<!-- -->

a)  Standardize continuous variables (e.g., hs and depres).

```{r}

```


b)  Fit a model to examine whether the effect of hate speech on depression symptoms depends on whether the participant has family support.

First, define appropriate priors for your model.

```{r}

```

Then, fit the model using the brm() function.

```{r}

```

Check for convergence issues using the WAMBS (When to Worry and How to Avoid the Misuse of Bayesian Statistics) checklist. Report whether it is safe to interpret the results.

```{r}

```


d)  Visualize the interaction between hate speech and family support.

```{r}

```


e)  Use hypothesis() or emtrends() to estimate and compare the effect of standardized hate speech (z_hs) for participants with and without family support.

```{r}

```

f)  Interpret the results
