---
title: "Final test"
format: html
---

#### Dataset and Model Setup

You will work with the following dataset and model. Carefully review the code and outputs before answering the questions that follow.

```{r}
library(tidyverse)
library(brms)
library(bayestestR)

inv_cloak <- read_csv("data_files/invisibility_cloak.csv")
inv_cloak <- inv_cloak %>% 
  mutate(Cloak = factor(Cloak, 0:1, c("no","yes")))
```

This dataset records mischievous behavior of individuals **with** and **without** an invisibility cloak. You will use Bayesian modeling to investigate whether having a cloak influences behavior.


---

### 1. Understanding Prior Distributions

```{r}
priors <- c(set_prior("normal(4, 2)", class = "b", coef = "Cloakno"),
            set_prior("normal(4, 2)", class = "b", coef = "Cloakyes"),
            set_prior("normal(0, 1)", class = "sigma"),
            set_prior("normal(0, 25)", class = "nu"))

validate_prior(Mischief ~ 0 + Cloak,
               data = inv_cloak,
               family = student,
               prior = priors)
```

**Questions:**

1.1. Explain the meaning of the prior `normal(4, 2)` set on both `Cloakno` and `Cloakyes`. Why might the researcher choose the same prior for both?

1.2. What is the interpretation of setting a prior on `sigma` and `nu` in a Student's t distribution?

1.3. Suppose you wanted to express more uncertainty about the `Cloakyes` effect. How would you change its prior?

---

### 2. Understanding the `brms` Model

```{r}
fit <- brm(Mischief ~ 0 + Cloak,
           data = inv_cloak,
           family = student,
           prior = priors,
           sample_prior = "yes")
```

**Questions:**

2.1. Why is the model specified as `Mischief ~ 0 + Cloak` instead of the more common `~ Cloak`?

2.2. What is the purpose of setting `sample_prior = "yes"`?

2.3. Why might a Student’s t distribution (`family = student`) be preferred over a Gaussian in this model?

---

### 3. Understanding Convergence Diagnostics

```{r}
bayesplot::mcmc_trace(fit)
```

```{r}
rhat(fit)
```

```{r}
neff_ratio(fit)
```

**Questions:**

3.1. What are you looking for in an MCMC trace plot to determine whether convergence has occurred?

3.2. What is the acceptable range for R-hat values? What do the values in this example suggest?

3.3. What does the `neff_ratio` tell you about the quality of the posterior samples?

---

### 4. Understanding Credible Intervals

```{r}
bayestestR::eti(fit)
```

**Questions:**

4.1. What can we conclude based on the values for `b_Cloakyes`?

4.2. How would changing the width of the ETI (e.g., to 90%) affect your conclusions?

---

### 5. Understanding Probability of Direction (PD)

```{r}
hypothesis(fit, "Cloakyes - Cloakno = 0") %>%
  .$samples %>%
  p_direction()
```

**Questions:**

5.1. What does the probability of direction measure in Bayesian terms?

5.2. What can we infer about the effect of wearing a cloak, based on the output of probability of direction?


---

### 6. Understanding Equivalence Testing

```{r}
hypothesis(fit, "Cloakyes - Cloakno = 0") %>%
  .$samples %>%
  equivalence_test(range = c(-1, 1))
```

**Questions:**

6.1. What is the purpose of a region of practical equivalence (ROPE) in Bayesian inference?

6.2. What does the value of the posterior distribution that lies inside the ROPE of (-1, 1) suggest?

---

### 7. Understanding Bayes Factors

```{r}
hypothesis(fit, "Cloakyes - Cloakno = 0")
```

**Questions:**

7.1. What is a Bayes factor and how is it interpreted?

7.2. What does the obtained value of Bayes factor suggests?



